---
title: "Reproducible Research:Peer Assessment 1"
output: html_document
---
It is assumed that the data to be analysed is already in the Directory.</br>
Packages Used in the analysis:
```{r first,warning=FALSE,message=FALSE,results='asis'}
library(knitr)
library(ggplot2)
library(dplyr)
```

<H2>Loading and preprocessing the data:</H2>
The orignal data is transformed into data_needed which will be needed for further analysis.

```{r}
unzip("activity.zip")
data=read.csv("activity.csv")
data$date=as.Date(data$date, format = "%Y-%m-%d")
data_needed=data[!(is.na(data$steps)),]
rownames(data_needed)=1:nrow(data_needed)
head(data_needed)
tail(data_needed)
```

<H2>What is mean total number of steps taken per day?</H2>
The data is grouped together by date.
```{r second}
data_group_date=group_by(data_needed,date)
ggplot(data_group_date, aes(date, steps)) + geom_bar(stat = "identity", colour = "#003399", fill = "#003399", width = 0.7) +  labs(title = "Histogram of Total Number of Steps Taken Each Day", x = "Date", y = "Total number of steps")
summarise(data_group_date, mean(steps),median(steps))
```

<H2>What is the average daily activity pattern?</H2>
The data is now grouped by intervals.
```{r third}
data_group_interval=group_by(data_needed,interval)
hist_data_interval=summarise(data_group_interval,mean(steps))
plot(hist_data_interval,type='l',ylab="Average Number of Steps",main="Average Daily Pattern",col="red")
```

```{r interval,echo=FALSE}
colnames(hist_data_interval)=c("interval","steps_mean")
row_max=which.max(hist_data_interval$steps_mean)
ans=as.character(hist_data_interval[row_max,1])
```
</br>The 5 minute interval which contains the maximum steps is <b>`r ans`</b>

<H2>Imputing missing values</H2>
```{r missing,echo=FALSE}
num_missing=sum(is.na(data))
```
The number of missing values are <b>`r num_missing`</b></br>
Now we are going to fill the intervals which have missing values by their interval means calculated over all the days.
```{r}
newData = data 
for (i in 1:nrow(newData)) {
  if (is.na(newData$steps[i])) {
    newData$steps[i] = hist_data_interval[which(newData$interval[i] == hist_data_interval$interval), ]$steps_mean
  }
}
```
</br>Now we will make a histogram.
```{r}
newdata_group_date=group_by(newData,date)
ggplot(newdata_group_date, aes(date, steps)) + geom_bar(stat = "identity", colour = "#003399", fill = "#003399", width = 0.7) +  labs(title = "Histogram of Total Number of Steps Taken Each Day", x = "Date", y = "Total number of steps")
```
</br>
Let us have a look at the mean and median.
```{r,echo=FALSE}
summarise(newdata_group_date, mean(steps),median(steps))
```
We see that that mean and median remains the same for previous entries but both increase for the new values added.

<H2>Are there differences in activity patterns between weekdays and weekends?</H2>
Function for calculating weekday or weekend.
```{r}
newData$typedate=ifelse(as.POSIXlt(newData$date)$wday %in% c(0,6), 'weekend', 'weekday')

```
Now let us make a panel plot of 5 minute intervals for weekdays and weekends.

```{r}
averageddata <- aggregate(steps ~ interval + typedate, data=newData, mean)
ggplot(averageddata, aes(interval, steps)) +   geom_line() +   facet_grid(typedate ~ .)  +labs(title="Weekday vs Weekend Average Steps taken",y="avarage number of steps",x="5-minute interval")
```